---
title: "EDA of Different Rates"
author: "Luke Benz"
date: "7/25/2020"
output: html_document
---

# Red and Yellow Cards

Have played around w/ $k-$game rolling averages of red and yellow cards per game with various values of $k$ (like 50, 100, 200). Open question to me is what is best value of $k$ to use here. (Note these are right windowed so average for first 100 games gets assigned to the point $x$ = 100). 


* Germany 1 + 2, Portugese Liga, Spain (Home YC up, Away YC down)
* Spain 1 (Away YC down)
* England 1 + 2, Italy 1 + 2 (Cards down in general)
* Sweden, Norway (Cards up in general)


__Open Questions/Notes:__ 

* How do we best attribute if rate is going up due to no fans vs continuing a trend of going as it had been before? 
* Should these rolling averages include games on both sides of the restart line?
* Optimal $k$ for rolling averages?
* Somewhere in this time period VAR was adopted by different leagues which I believe allows red cards to be reviewed (and non-red card plays to be reviewed for a red card)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse)
library(zoo)
theme_set(theme_bw() +
            theme(plot.title = element_text(hjust = 0.5, size = 24),
                  plot.subtitle = element_text(hjust = 0.5, size = 18),
                  axis.title = element_text(size = 20),
                  strip.text = element_text(size = 14),
                  legend.position = "bottom"))
```


```{r, fig.height = 9/1.2, fig.width = 16/1.2}
league_info <- read_csv("../league_info.csv")
k <- 100
for(i in 1:nrow(league_info)) {
  league <- league_info$alias[i]
  file_league <- gsub("\\s", "_", tolower(league))
  files <- dir(paste0("../fbref_data/", file_league), full.names = T)
  
  game_stats <- 
    map_dfr(files, read_csv) %>% 
    arrange(date) %>% 
    ### A few games for which stats not available
    filter(!is.na(away_red_cards), !is.na(home_red_cards),
           !is.na(away_yellow_cards), !is.na(home_yellow_cards))
  
  
  
  game_stats <- 
    game_stats %>% 
    mutate("rolling_home_rcpg" = c(rep(NA, k-1), rollmean(home_red_cards, k, align = "right")),
           "rolling_home_ycpg" = c(rep(NA, k-1), rollmean(home_yellow_cards, k, align = "right")),
           "rolling_away_rcpg" = c(rep(NA, k-1), rollmean(away_red_cards, k, align = "right")),
           "rolling_away_ycpg" = c(rep(NA, k-1), rollmean(away_yellow_cards, k, align = "right")),
           "game_id" = 1:nrow(.))
  
  card_stats <- 
    game_stats %>% 
    select(date, game_id, contains("rolling")) %>% 
    pivot_longer(cols = contains("rolling"),
                 names_to = "stat",
                 values_to = "cards_per_game") %>% 
    mutate("card_type" = ifelse(grepl("rcpg", stat), "Red Cards", "Yellow Cards"),
           "team" = ifelse(grepl("home", stat), "Home", "Away")) 
  
  restart_date <- as.Date(league_info$restart_date[league_info$alias == league], "%m/%d/%y")
  restart_ix <- min(game_stats$game_id[game_stats$date == restart_date])
  
  p <- 
    ggplot(card_stats, aes(x = game_id, y = cards_per_game)) +
    geom_vline(xintercept = restart_ix, lty = 2) +
    facet_wrap(~card_type, scales = "free_y") +
    geom_line(aes(color = team)) +
    labs(x = paste("Games Since Start of", min(game_stats$season), "Season"),
         y = "Cards/Game",
         color = "",
         title = paste("Rolling", k, "Game Average of Cards Assessed Per Game"),
         subtitle = paste(league, "Since Start of", min(game_stats$season), "Season"))
  
  print(p)
  
}

```

# Goals

Same as above w/ home and away goals.

__Notes and Observations:__

* German Bundesliga first time home goals per game ever $\leq$ away goals per game during this time period.
* Bundeliga 2. scoring up all around
* Home GPG Down a lot in Austria. Similar trend in Spain 1.
* Portugal home GPG up and Away GPG down after being at home GPG at "record" low for the time period
* Sharp home GPG drop in Greece. This had been country amongst this group w/ largest home field avantage (see https://github.com/lbenz730/soccer_hfa/blob/master/simulation_figures/pre_covid_hfa.png)
* Scoring down in Sweden, up in Norway and Italy 1
* Widening of the gap in Switzerland and Italy 2 (two of the leagues w/ lowest pre-covid HFA)




```{r, fig.height = 9/1.2, fig.width = 16/1.2}
league_info <- read_csv("../league_info.csv")
k <- 100
for(i in 1:nrow(league_info)) {
  league <- league_info$alias[i]
  file_league <- gsub("\\s", "_", tolower(league))
  files <- dir(paste0("../fbref_data/", file_league), full.names = T)
  
  game_stats <- 
    map_dfr(files, read_csv) %>% 
    arrange(date) %>% 
    ### A few games for which stats not available
    filter(!is.na(home_score), !is.na(away_score))
  
  
  game_stats <- 
    game_stats %>% 
    mutate("rolling_home_gpg" = c(rep(NA, k-1), rollmean(home_score, k, align = "right")),
           "rolling_away_gpg" = c(rep(NA, k-1), rollmean(away_score, k, align = "right")),
           "game_id" = 1:nrow(.))
  
  goal_stats <- 
    game_stats %>% 
    select(date, game_id, contains("rolling")) %>% 
    pivot_longer(cols = contains("rolling"),
                 names_to = "stat",
                 values_to = "goals_per_game") %>% 
    mutate("team" = ifelse(grepl("home", stat), "Home", "Away")) 
  
  restart_date <- as.Date(league_info$restart_date[league_info$alias == league], "%m/%d/%y")
  restart_ix <- min(game_stats$game_id[game_stats$date == restart_date])
  
  p <- 
    ggplot(goal_stats, aes(x = game_id, y = goals_per_game)) +
    geom_vline(xintercept = restart_ix, lty = 2) +
    geom_line(aes(color = team)) +
    labs(x = paste("Games Since Start of", min(game_stats$season), "Season"),
         y = "Goals/Game",
         color = "",
         title = paste("Rolling", k, "Game Average of Goals Per Game"),
         subtitle = paste(league, "Since Start of", min(game_stats$season), "Season"))
  
  print(p)
  
}

```

# Fouls

__Notes and Obervations:__

* A lot of these foul numbers are bad for smaller leagues. Either missing first few seasons or just 0 for all games which we know can't be the case. If we use fouls/game, we probably we either remove games where both home team and away team commit 0 fouls, or we just focus on the big leagues.
* German Bundesliga home fouls/game up compared to recent trends but hard to say if this is not some reversion to where it usually is. I'd say less likely given what we see in terms of YC/game. Similar trend in Italy.
* England has seen maybe slight uptick in fouls for home team but this goes back to the question of is this less HFA or is is continution of seemingly longer recent trend.
* Spain, maybe fouls down a little across the board.
* Austria away fouls/game down.


```{r, fig.height = 9/1.2, fig.width = 16/1.2}
league_info <- read_csv("../league_info.csv")
k <- 100
for(i in 1:nrow(league_info)) {
  league <- league_info$alias[i]
  file_league <- gsub("\\s", "_", tolower(league))
  files <- dir(paste0("../fbref_data/", file_league), full.names = T)
  
  game_stats <- 
    map_dfr(files, read_csv) %>% 
    arrange(date) %>% 
    ### A few games for which stats not available
    filter(!is.na(home_fouls_committed), !is.na(away_fouls_committed))
  
  
  game_stats <- 
    game_stats %>% 
    mutate("rolling_home_fpg" = c(rep(NA, k-1), rollmean(home_fouls_committed, k, align = "right")),
           "rolling_away_fpg" = c(rep(NA, k-1), rollmean(away_fouls_committed, k, align = "right")),
           "game_id" = 1:nrow(.))
  
  foul_stats <- 
    game_stats %>% 
    select(date, game_id, contains("rolling")) %>% 
    pivot_longer(cols = contains("rolling"),
                 names_to = "stat",
                 values_to = "fouls_per_game") %>% 
    mutate("team" = ifelse(grepl("home", stat), "Home", "Away")) 
  
  restart_date <- as.Date(league_info$restart_date[league_info$alias == league], "%m/%d/%y")
  restart_ix <- min(game_stats$game_id[game_stats$date == restart_date])
  
  p <- 
    ggplot(foul_stats, aes(x = game_id, y = fouls_per_game)) +
    geom_vline(xintercept = restart_ix, lty = 2) +
    geom_line(aes(color = team)) +
    labs(x = paste("Games Since Start of", min(game_stats$season), "Season"),
         y = "Fouls Committted/Game",
         color = "",
         title = paste("Rolling", k, "Game Average of Fouls Committed Per Game"),
         subtitle = paste(league, "Since Start of", min(game_stats$season), "Season"))
  
  print(p)
  
}

```

# Penalties

__Notes and Obervations:__

* Away penalties up a lot in Greece (again, this is country in our dataset w/ highest pre-covid home field advantage).
* Away penalties at a new 5 year high in Spain 1. (Same for Turkey + Russia).
* Penalties up across the board in Italy 1 + Norway


```{r, fig.height = 9/1.2, fig.width = 16/1.2}
league_info <- read_csv("../league_info.csv")
k <- 100
for(i in 1:nrow(league_info)) {
  league <- league_info$alias[i]
  file_league <- gsub("\\s", "_", tolower(league))
  files <- dir(paste0("../fbref_data/", file_league), full.names = T)
  
  game_stats <- 
    map_dfr(files, read_csv) %>% 
    arrange(date) %>% 
    ### A few games for which stats not available
    filter(!is.na(home_pk_attempts), !is.na(away_pk_attempts))
  
  
  game_stats <- 
    game_stats %>% 
    mutate("rolling_home_pkpg" = c(rep(NA, k-1), rollmean(home_pk_attempts, k, align = "right")),
           "rolling_away_pkpg" = c(rep(NA, k-1), rollmean(away_pk_attempts, k, align = "right")),
           "game_id" = 1:nrow(.))
  
  pk_stats <- 
    game_stats %>% 
    select(date, game_id, contains("rolling")) %>% 
    pivot_longer(cols = contains("rolling"),
                 names_to = "stat",
                 values_to = "pk_per_game") %>% 
    mutate("team" = ifelse(grepl("home", stat), "Home", "Away")) 
  
  restart_date <- as.Date(league_info$restart_date[league_info$alias == league], "%m/%d/%y")
  restart_ix <- min(game_stats$game_id[game_stats$date == restart_date])
  
  p <- 
    ggplot(pk_stats, aes(x = game_id, y = pk_per_game)) +
    geom_vline(xintercept = restart_ix, lty = 2) +
    geom_line(aes(color = team)) +
    labs(x = paste("Games Since Start of", min(game_stats$season), "Season"),
         y = "Penalties/Game",
         color = "",
         title = paste("Rolling", k, "Game Average of Penalty Kicks Per Game"),
         subtitle = paste(league, "Since Start of", min(game_stats$season), "Season"))
  
  print(p)
  
}

```


### Cards/Foul: What Percentage of Fouls are Yellow Cards
```{r, fig.height = 9/1.2, fig.width = 16/1.2}
league_info <- read_csv("../league_info.csv")
k <- 100
for(i in 1:nrow(league_info)) {
  league <- league_info$alias[i]
  file_league <- gsub("\\s", "_", tolower(league))
  files <- dir(paste0("../fbref_data/", file_league), full.names = T)
  
  game_stats <- 
    map_dfr(files, read_csv) %>% 
    arrange(date) %>% 
    ### A few games for which stats not available
    filter(!is.na(home_fouls_committed), 
           !is.na(away_fouls_committed),
           !is.na(away_yellow_cards), 
           !is.na(home_yellow_cards),
           ### Small # of games removed since they must have bad fouls
           home_yellow_cards <= home_fouls_committed,
           away_yellow_cards <= away_fouls_committed) %>% 
    mutate("home_yc_rate" = home_yellow_cards/(home_fouls_committed + 0.0000001),
           "away_yc_rate" = away_yellow_cards/(away_fouls_committed + 0.0000001))
  
  
  if(nrow(game_stats) > 100) {
    game_stats <- 
      game_stats %>% 
      mutate("rolling_home_yc_rate" = c(rep(NA, k-1), rollmean(home_yc_rate, k, align = "right")),
             "rolling_away_yc_rate" = c(rep(NA, k-1), rollmean(away_yc_rate, k, align = "right")),
             "game_id" = 1:nrow(.))
    
    yc_rate_stats <- 
      game_stats %>% 
      select(date, game_id, contains("rolling")) %>% 
      pivot_longer(cols = contains("rolling"),
                   names_to = "stat",
                   values_to = "yc_rate") %>% 
      mutate("team" = ifelse(grepl("home", stat), "Home", "Away")) 
    
    restart_date <- as.Date(league_info$restart_date[league_info$alias == league], "%m/%d/%y")
    restart_ix <- min(game_stats$game_id[game_stats$date == restart_date])
    
    p <- 
      ggplot(yc_rate_stats, aes(x = game_id, y = yc_rate)) +
      geom_vline(xintercept = restart_ix, lty = 2) +
      geom_line(aes(color = team)) +
      labs(x = paste("Games Since Start of", min(game_stats$season), "Season"),
           y = "% of Fouls that are\nYellow Cards",
           color = "",
           title = paste("Rolling", k, "Game Average of Yellow Cards/Fouls Rate"),
           subtitle = paste(league, "Since Start of", min(game_stats$season), "Season")) +
      scale_y_continuous(labels = scales::percent)
    
    print(p)
  }
  
}
```
```