---
title: "Data Coverage"
author: "Luke Benz"
date: "7/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r}
### Helper Functions to Get Summary Stats
league_summary <- function(league) {
  file_league <- gsub("\\s", "_", tolower(league))
  files <- dir(paste0("../fbref_data/", file_league), full.names = T)
  
  x <- map_dfr(files, read_csv)
  
  df_stats <- 
    map_dfr(unique(x$season), ~sum_stats(filter(x, season == .x))) %>% 
    mutate("league" = league)
  
  return(df_stats)
}

sum_stats <- function(df) {
  enframe(apply(df, 2, function(x) !all(is.na(x)))) %>%
    filter(!name %in% c("date", "home", "away")) %>% 
    rename("variable" = name,
           "present" = value) %>% 
    mutate("season" = as.character(df$season[1]))
}
```


```{r}
league_info <- read_csv("../league_info.csv")
data_summary <- map_dfr(league_info$alias, league_summary)
```

# Coverage
We make the assumption if a variable is present, it is present since that year (i.e. they only add more data over time -- not take it away). This can be reasonablly checked based on what we decided to end up using but it creates a simplified heuristic for variable selection/better understanding the data. 

```{r}
data_summary <- 
  data_summary %>% 
  filter(present) %>% 
  group_by(league, variable) %>% 
  summarise("since" = min(season)) %>% 
  pivot_wider(names_from = variable, 
              values_from = since)

data_summary[is.na(data_summary)] <- as.character(emo::ji("x"))

kable(data_summary, escape = F, align = "ccccccc") %>%
  kable_styling("striped", full_width = F, position = "center") %>%
  row_spec(0, bold = T, font_size = 12) %>% 
  add_header_above(c("Data Coverage Summary" = ncol(data_summary)), bold = T, font_size = 24)
```


# Referees
Referees look surprisingly clean! Few cases in smaller leagues that could possibly need cleaning (Magnus Lindgren vs. Marcus Lindgren?) but for the big leagues it looks fine.
```{r}
df_ref <- NULL
for(i in 1:nrow(league_info)) {
  league <- league_info$alias[i]
  file_league <- gsub("\\s", "_", tolower(league))
  files <- dir(paste0("../fbref_data/", file_league), full.names = T)
  x <- map_dfr(files, read_csv)
  
  df_ref <- 
    enframe(table(x$referee), name = "Referee Name", value = "# of Observations") %>% 
    mutate("league" = league) %>% 
    bind_rows(df_ref)
}

df_ref %>%
  arrange(`Referee Name`) %>% 
  kable() %>% 
  kable_styling("striped", full_width = F, position = "center") %>%
  row_spec(0, bold = T, font_size = 12) %>% 
  add_header_above(c("Referee Summary" = 3), bold = T, font_size = 24) 

```
